---
title: "R Learner's Permit Exam"
author: "Andrew J. Greenlee, Ph.D."
date: "2/19/2021"
output:
  html_notebook:
    toc: yes
---

# Introduction
Over the course of the first few weeks of our class, you have learned some basic elements of the R programming language and RStudio IDE. As we move forward with our class, we will continue to add knowledge of new packages, tools, and data within R.

This R Learner's Permit is designed to assess your knowledge of the core elements of the R language and software that we have focused on so far in class.

## Instructions

Please follow the instructions contained in this R Notebook, which involve filling in code and generating output. Most instructions ask you to add or fill in code chunks. Others ask you to provide a written interpretation in the notebook portion of the document.

Complete as many items as you can. If you run into trouble completing an item, add comments to your code or in the notebook describing where you are running into problems, and what you think the next step might be to solving the problem you're having.

Complete as many items as you can and then upload your R Notebook to the "R Learners Permit" folder in your assignment repository on GitHub.

## The Data

In the "data" subfolder of this project, you will find one .rds file called NYC_Violations.rds. This is the data you will use to earn your learner's permit. These are "real world" data derived from two sources:

1. New York City [housing maintainance code violation](https://data.cityofnewyork.us/Housing-Development/Housing-Maintenance-Code-Violations/wvxf-dwi5) data for the years 2009 to 2019
2. Data on property characteristics from the [Primary Land Use Tax Lot Output (PLUTO)](https://www1.nyc.gov/site/planning/data-maps/open-data/dwn-pluto-mappluto.page) database

I have merged selected columns from these datasets for you (you will find data documentation and metadata for both datasets in the documentation subfolder), and after taking a close look at the dataset and the documentation, you should be able to identify which portions of the data you will work with come from each dataset.

The PLUTO data contains information on each [tax lot](http://gis.nyc.gov/taxmap/map.htm) within the city of New York. By merging these data which contain characteristics of buildings and land with the building code violation data, the data you have to work with describes the characteristics of the building and property where the violation occurred as well as the nature of the violation. It is essential for you to spend the time to think carefully about what each row in the building code violation data represents.

# Earn Your Learner's Permit

With that background in mind, its time to earn your learner's permit.

## Task 1: Load and Describe the Data

At the beginning of the script, we will create a holding area within which you can load any packages you may need to work with as you year your R Learner's Permit (if you are not using any packages, please leave this area blank):
```{r, echo=FALSE}
# Your Code Here
library(tidyverse)
library(ggplot2)
```

Use the `readRDS()` command to load the data into an object with the name of your choice.
```{r}
# Your Code Here
violationsNYC <- readRDS("C:data/NYC_Violations_WB.rds")
```

Describe the structure of the data, including the number of rows and columns in the data. Create a table outside of the code chunk below in your notebook that describes each variable name and what each row in the data represents (a [Markdown table generator](https://www.tablesgenerator.com/markdown_tables) may be useful):
```{r}
# Your Code Here
nrow(violationsNYC)
ncol(violationsNYC)
str(violationsNYC)
```

**Your Narrative Description of the Data Structure:**
Note: I decided not to use a markdown table because it looked too crowded with the lines to me 

This dataset has 12 columns and 4,451,225 rows. Each row represents one violation and with it comes information in each of the 12 columns. The following table shows the title, data type, and description of each column:

Column Name: 
BBL              chr    Unique ID for each tax lot                 
Borough          chr    Location in New York City's 5 Boroughs
NTA              chr    Location in neighborhood
ViolationStatus  chr    Whether the case is open or closed
InspectionYear   num    The year the site was inspected
unitstotal       num    The total number of units at each site
ownername        chr    The name of the owner of the property
assessland       num    The assessed value of the land
assesstot        num    The assessed total value of the property
yeabuilt         num    The year the building was built
histdist         chr    The historic district the property is located in
OrderNumber      num    The code for the violation

***
## Task 2: Summarize Violations

Describe the number of unique violations for each year (information on the year the unit was inspected for violations is contained in the InspectionYear variable):
```{r}
# Your Code Here
violationsYear <- violationsNYC %>% count(InspectionYear)
```

Now, build upon your summary code to make a plot (geometry of your choice) of inspections by year:
```{r}
ggplot(violationsYear, aes(InspectionYear, n)) +
  geom_col(aes(fill = InspectionYear)) +
  labs(x = "Year of Inspection", y = "Number of Inspections", title =  "Property Violations by Year in New York City")
```

In your own words, what is the overall trend with regards to the number of violations issued between 2009 and 2019?:

**Your Narrative Description:**

The general trend is that the number of violations increases as the years go on. 2019 has the greatest number of violations and 2009 has the smallest number. I used the gradient color to show that increase from 2009 to 2019.

Next, describe the number of unique violations for each New York City borough. Do this using both *base* r code and using `dplyr` code that utilizes the `summarise()` function:
```{r}
# Your Code Here
# Using count
violationsBorough <- violationsNYC %>% count(Borough) %>% rename(Violations = n)

#Using summarise 
violationsNYC %>% group_by(Borough) %>% summarise(n = n())
```
***
## Task 3: More Detail on Violations

Let's create some more detailed summaries of violations. 

Create a table that enumerates by borough the number of violations that are currently closed or open (ViolationStatus). Remove the `NA` values so that you are just reporting open and closed cases:
```{r}
# Your Code Here
violationsNYC <- violationsNYC %>% filter(!is.na(ViolationStatus))
violationsBoroughStatus <- violationsNYC %>% count(Borough, ViolationStatus) %>% group_by(ViolationStatus)
```
See if you can refine the formatting of this table. Although we haven't formally introduced `pivot_wider()` yet, try using it so that you produce a summary in which there are three columns - the borough, the number of closed cases, and the number of open cases (remember you can search for documentation, e.g. `?pivot_wider()`). You will need to supply pivot_wider with an argument for `names_from` and an argument for `values_from`. Names should come from the Violation Status column, and values should come from the column in which the count of violations is contained.
```{r}
# Your Code Here
violationsBoroughStatus <- violationsBoroughStatus %>% pivot_wider(names_from = ViolationStatus, values_from = n)
```
Now create a chart (geom of your choice) that represents the number of *open cases* by borough by the year in which the violation occurred:
```{r}
ggplot(violationsBoroughStatus, aes(Borough, Open, fill = Borough)) + 
  geom_col() +
  labs(y = "Number of Open Cases", title =  "Open Violation Cases by Borough in New York City")
```

Now, create a table with the total number of *buildings* with violations (remember, there may be multiple violations per building in the data). Remember that the BBL column contains a unique identification number for each building (technically it's each tax lot, but we'll use that as a proxy for building). You can use the `distinct()` command to create a subset of the distinct buildings with violations. You could also use the `summarise()` command to accomplish something similar...

```{r}
# Your Code Here
?distinct
violationsUnique <- violationsNYC %>% distinct(BBL, .keep_all = TRUE)
```

Let's create a final summary table of buildings with violations and total violations by borough. Use `dplyr` [join syntax](https://dplyr.tidyverse.org/reference/join.html) to accomplish this. Your table should have a column containing the borough name, a column containing the total number of buildings with violations, and a column containing the total number of violations.
```{r}
# Your Work Here
violationsUnique <- violationsUnique %>% count(Borough) %>% rename(Buildings = n)
buildingBorough <- left_join(violationsUnique, violationsBorough, by = "Borough")
```

Finally, add to this table, the average violations per building by borough. You can use the `mutate()` command to add this column to your existing table.

```{r}
# Your Work Here
buildingBorough <- buildingBorough %>% mutate(Avg_per_Building = Violations/Buildings)
```
Which boroughs contain higher rates of violations per building?

**Your Narrative Description on Violations by Borough:**

The Bronx has a significantly higher rate of violations per building and Manhattan is the next highest.

***

## Task 4: Building Ages

Create a table that displays the mean and median ages (yearbuilt) of buildings with violations by borough. Remember that individual buildings may have multiple observations - you will need to figure out how to create a summary of the dataset that includes one record per building to generate an unbiased summary.
```{r}
# Your Code Here
agesUnique <- violationsNYC %>% distinct(BBL, .keep_all = TRUE) %>% mutate(Age = 2021 - yearbuilt, na.rm = TRUE) %>% group_by(Borough) %>% summarise_at(vars(Age), funs(mean(., na.rm = TRUE), median(., na.rm = TRUE)))
```

Add to the table that you just made two columns that include the ages of the oldest and youngest buildings for each borough (since the years are numeric, the functions `min()` and `max()` may be useful here):
```{r}
# Your Code Here
violationsNYC <- violationsNYC %>% filter(!is.na(yearbuilt))
buildingAges <- violationsNYC %>% distinct(BBL, .keep_all = TRUE) %>% mutate(Age = 2021 - yearbuilt) %>% group_by(Borough) %>% summarise(MaxAge = max(Age, na.rm = TRUE), MinAge = min(Age, na.rm = TRUE)) %>% arrange(Borough)

```

What do you discover when trying to do this, and how might you modify your code to rectify it?

**Your Narrative Description of this Table and code to address your discovery:**

I realized when doing this that the yearbuilt column has 0's in it. I removed missing data but now I need to remove the 0 years as well. 

```{r}
# Your Code here
buildingAges <- violationsNYC %>% filter(yearbuilt != 0) %>% distinct(BBL, .keep_all = TRUE) %>% mutate(Age = 2021 - yearbuilt) %>% group_by(Borough) %>% summarise(MaxAge = max(Age, na.rm = TRUE), MinAge = min(Age, na.rm = TRUE)) %>% arrange(Borough)

```

**Your Narrative Description of the implications of any modifications you made between the last two code chunks:**

I removed 0's from the data in order to find the max and min but that does mean that for that particular analysis we aren't looking at every available data input.

***
## Task 5: Decade Comparison
Thinking about each borough, create a table that would allow us to compare the frequency of violations in 2009 with the frequency of violations in 2019. 

```{r}
# Your Code Here
violations0919 <- violationsNYC %>% count(Borough, InspectionYear) %>% filter(InspectionYear == "2009" | InspectionYear == "2019") %>% pivot_wider(names_from = InspectionYear, values_from = n)
violations0919
```

How has the frequency of violations changed over the decade?

**Your Narrative Description:**

The number of violations has increased significantly from 2009 to 2019 in every borough.

***

## Task 6: Frequency of Violations by Violation Type

The OrderNumber column contains information on the actual violation recorded (the housing maintenance code violation user guide contains a narrative description of each order number).

What is the most frequent code violation recorded in each borough?
```{r}
# Your Code Here
violationsType <- violationsNYC %>% count(Borough, OrderNumber) %>% group_by(Borough) %>% slice(which.max(n)) %>% arrange(Borough)
violationsType
```

**Your Narrative Description:**

Bronx           508     "REPAIR THE BROKEN OR DEFECTIVE PLASTERED SURFACES AND PAINT IN A UNIFORM COLOR"
Brooklyn        780     "OWNER FAILED TO FILE A VALID REGISTRATION STATEMENT WITH THE DEPARTMENT AS REQUIRED BY ADM CODE"
Manhattan       508     "REPAIR THE BROKEN OR DEFECTIVE PLASTERED SURFACES AND PAINT IN A UNIFORM COLOR
Queens          780     "OWNER FAILED TO FILE A VALID REGISTRATION STATEMENT WITH THE DEPARTMENT AS REQUIRED BY ADM CODE"
Staten Island   780     "OWNER FAILED TO FILE A VALID REGISTRATION STATEMENT WITH THE DEPARTMENT AS REQUIRED BY ADM CODE"

***
## Task 7: Building Owners with the Most Violations

Which identifiable owner (ownername) has the highest rate of violations per unit?
```{r}
# Your Code Here
violationsOwner <- violationsNYC %>% filter(unitstotal != 0) %>% count(ownername, OrderNumber, unitstotal) %>% group_by(ownername) %>% mutate(Rate_per_Unit = n/unitstotal) 

violationsOwner[order(-violationsOwner$Rate_per_Unit),]

```

**Your Narrative Description:**

The identifiable owner with the highest rate of violations per unit is the NYC Housing Authority.

***
## Task 8: Buildings in Historic Districts

For each Borough, how many violations occur in *buildings* built after 1957 which are located in historic districts (histdist)?

```{r}
# Your Code Here
newHistoric <- violationsUnique %>% filter(!is.na(histdist), yearbuilt > 1957)
newHistoricBorough <- newHistoric %>% count(Borough)
```

Add to this table the average assessed value of the land (assessland), the average assessed value of the buildings, and the average total assessed value.

```{r}
# Your Code Here
newHistoric <- newHistoric %>% group_by(Borough) %>% summarise(AvgLandValue = mean(assessland), AvgBuildValue =mean(assesstot - assessland), AvgTotValue = mean(assesstot)) %>% left_join(newHistoricBorough, newHistoric, by = "Borough")
```

For those buildings in historic districts, how different are the ratios of the assessed values for land and buildings in each borough? What is your interpretation of these differences?

```{r}
# Your Code Here

```
**Your Narrative Description:**

I'm honestly not sure what this question is asking but that is probably because my brain is fried after a long week of R work. 
I'll hopefully come back to this question later next week. This was a great lab, thank you!

***
# The End
That's it - you're done! Your permit will be assessed and returned shortly. Please submit your R notebook to the appropriate GitHub folder. Rename the notebook so that the name is lastname_firstname_Permit.Rmd (e.g. Greenlee_Andrew_Permit.Rmd). You just need to upload the .Rmd file - no need to include the whole file structure.